%https://drive.google.com/file/d/0B7QKeIdz_U4vcFBkb2M4ZzhRdHc/view
\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage[margin=3cm]{geometry}

\title{capgeek.com FAQ}
\author{Matthew Wuest}

\begin{document}

\maketitle

\begin{enumerate}
\item \textsc{Are the Flyers, with Chris Pronger, and the Bruins, with Marc Savard, circumventing the cap?}

\bigskip

Short answer, no. Longer answer follows.

First, both players have serious concussions that affected not only their hockey careers, but, according to a variety of reports, their everyday lives.

Second, the league’s standard player’s contract dictates that any player who is “disabled and unable to perform his duties as a hockey player by reason of an injury sustained during the course of his employment as a hockey player” is “entitled to receive his remaining” salary and signing bonuses until expiration of his contract. This means that although their injuries could indeed be classified as career-ending, it is their right within the CBA to delay retirement in order to collect the full amounts due to them each season until their contract expires. If they retired, they would forfeit those payments.

Third, since each player has an injury that will obviously keep them out for at least 10 games and 24 days, their respective teams can use long-term injured reserve to replace the cap amounts of each player.

It should be noted that retirement for Savard prior to expiration of his contract would result in recapture penalties to the Bruins, while retirement for Pronger prior to expiration of his contract would result in his full cap hit applying to the Flyers because of the “over-35” rule.

In a nutshell, the player benefits financially and the team benefits from a cap perspective — and in both cases, basic CBA regulations are being followed.

\textbf{CBA Reference:} Article 23.4, P. 151, 2013 CBA

\bigskip
\bigskip

\item
\textsc{Can a team buy out an injured player?}

\bigskip

The CBA technically does not allow injured players to be bought out. However, in rare situations, it has occurred — typically by mutual agreement between team and player.

The most prominent example occurred in 2011 with the New York Rangers and Chris Drury. The Rangers intended to buy Drury out of the final year of his contract for the 2011-12 season. On June 17, 2011, the New York Post reported the Rangers would be "unable to buy out the final season of Drury's contract because of a degenerative knee condition" that would have rendered him "medically unable to play" in 2011-12. But on June 29, 2011, the New York Post reported Drury decided "against applying for a medical exception" that — if he were "declared medically unable to perform" — would have allowed him to collect the full \$5 million due to him. Instead, as a result of Drury's decision, the Rangers were able to buy him out for approximately \$3.33 million with cap charges of approximately \$3.717 million in 2011-12 and \$1.667 million in 2012-13.

Another example of this occurred in 2008 with the Carolina Hurricanes and concussion-plagued David Tanabe. In June of that year, Carolina "sought to buy out" Tanabe but was thwarted when the "NHLPA filed a grievance based on the terms of the CBA," according to a report in the Raleigh News and Observer on Aug. 22, 2008. On Oct. 17, 2008, the News and Observer reported that the Hurricanes and Tanabe had "reached a settlement." Tanabe, who was due \$900,000 in the final season of his contract for the 2011-12 season, instead received "\$850,000 over the next three years." It was not made public how Tanabe's contract counted against the salary cap, although it is likely the standard buyout formula applied.

The other example of interest also occurred in 2008 with the Boston Bruins and Glen Murray. In June of that year, Murray was bought out of the final year of his contract — allowing him to receive approximately \$2.76 million of the \$4.15 million he was due — in June 2008. According to a report on TSN.ca on Nov. 5, 2008, Murray required ankle surgery on Nov. 11, 2011, and although he did not realize it at the time of the buyout, believed the injury was "pre-existing and should have prohibited the Bruins from buying him out." However, the Boston Globe reported on June 18, 2009, that "arbitrator Richard Bloch ruled that the Bruins were within their rights (in 2008) to buy out the final year of Murray's deal." According to the report, Murray had argued "he was injured at the time the Bruins rendered the buyout, and thus he should have been ineligible to be bought out, per the protections of the Collective Bargaining Agreement. Bloch disagreed, noting in his written decision that Murray was too late seeking a second medical opinion, a key and necessary ingredient in structuring his grievance."

\bigskip
\bigskip
\item
\textsc{How are player ages calculated?}

\bigskip

Age is calculated different ways in the NHL-NHLPA collective bargaining agreement.

\smallskip

\textit{Entry-Level Signing Age}

\smallskip

A player’s entry-level signing age — used for determining length of the contract and slide eligibility — is calculated on Sept. 15 of the calendar year the contract was signed, per Section 9.2 of the CBA. However, per Section 9.1 (d) (i) of the CBA, players turning 20 between Sept. 16 and Dec. 31 in the calendar year the contract was signed are an exception to the rule. Such players are deemed to be 20 years of age rather than 19.

\smallskip

\textit{Waiver Signing Age}

\smallskip

A player’s waiver signing age — used for determining waiver eligibility — is outlined in Section 13.4 of the CBA and is calculated as follows:
18 — Player must turn 18 between Jan. 1 and Sept. 15 in the entry-draft calendar year preceding the first season of his first contract.
19+ — The age the player reaches in the entry-draft calendar year preceding the first season of his first contract.

\bigskip
\bigskip
\item

\textsc{How can Long-Term Injured Reserve cause a bonus overage?}

\bigskip

Teams requiring long-term injury relief to fit a roster under the salary cap throughout the season can be at risk of ending up with a bonus overage. That's because utilizing LTIR results in the team finishing the season with \$0 cap space. For teams in these positions with players earning performance bonuses, those bonuses are tacked onto their final cap payroll as an overage and carried forward as a penalty the following season.

\textbf{CBA Reference:} Section 50.10 (d) (P. 226-231)

\bigskip
\bigskip
\item

\textsc{How can teams receive cap relief during roster emergencies?}

\bigskip

Article 50.10 (e) of the new collective bargaining agreement includes a provision for cap-exempt recalls during roster emergencies in which injuries and/or suspensions result in insufficient cap room and a shortage of skaters for more than one game.

The following provisions apply:

\begin{itemize}
\item Cap room must be less than the minimum salary plus \$100,000, ie. \$650,000 in 2013-14.

\item The club must be unable to add a player with a cap hit equal to the minimum NHL salary plus \$100,000, ie. \$650,000 in 2013-14, using long-term injury exception.

\item The team must have fewer than 18 skaters and two goaltenders on its roster for at least one game before the recall(s) can occur.

\item The recalled player(s) must each earn less than the NHL's minimum salary plus \$100,000, ie. \$650,000 in 2013-14, and can only remain on the roster during the roster-emergency period.

\end{itemize}


\textbf{CBA Reference:} Article 50.10 (e) (P. 294-295)

Link: \href{https://web.archive.org/web/20141007101150/http://www.theglobeandmail.com/try-it-now/?articleId=14545832}{New CBA Twist Could Help Teams with Cap Crunch} by James Mirtle, Globe and Mail, Sept. 26, 2013

\bigskip
\bigskip
\item
\textsc{How do buyouts work?}

\bigskip

Please note, this FAQ entry refers to ordinary-course buyouts that count against the cap. For details on compliance buyouts that do not count against the cap, please read our separate FAQ entry titled "How do compliance buyouts work?"

Teams are entitled to buy out player contracts for a portion of the remaining value of the contract — paid over a period of twice the remaining length of the contract. Following are the buyout amounts:

\begin{itemize}

\item Younger than age 26 at the time of buyout, 1/3 the remaining value

\item Age 26 or older at the time of the buyout, 2/3 the remaining value

\end{itemize}

When a player is bought out, the team still takes a cap hit for the player over a period of twice the remaining length of the contract. The amount of the cap hit (by year) is determined as follows:

\begin{enumerate}

\item Take the actual salary due for each remaining year
\item Take the Averaged Player Salary (cap hit) for the current contract
\item Calculate the buyout amount (as described above)
\item Spread the buy-out amount evenly over twice the remaining years of the contract
\item Take the number in No. 1 and subtract the number in No. 4. This is the "buyout savings."
\item Take the cap hit from No. 2 and subtract the buyout savings from No. 5.

\end{enumerate}

This calculation is done for each year, meaning the buyout cap hit is not necessarily the same for all years. It can even be negative, meaning a team can receive a credit.

This formula also applies to 35-plus contracts. (See Also: What's a 35-plus contract?)

While compliance buyouts prohibit a player from rejoining the team that bought him out for one season, there is no rule prohibiting this with ordinary-course buyouts.

In order for a team to execute a buyout, the player must first clear unconditional waivers. A waiver claim by another team pre-empts the buyout process. Players with no-move clauses can reject the option of waivers and proceed immediately to a buyout.

The buyout period begins the later of June 15 or 48 hours after the Stanley Cup Final ends. It concludes on June 30 at 5 p.m. ET. Teams with salary arbitration filings can additionally execute buyouts during a 48-hour period beginning on the third day following the team's last award/settlement. The following notes pertain only to the second buyout window following salary arbitration: a) the second buyout period does not apply in the event a team has only one club-elected salary arbitration case; b) buyouts during this second window are limited to three per team over the term of the CBA; and c) players bought out during this window had to be on the team's reserve list as of the most recent trade deadline and must have a cap hit of \$2.75 million or more.

Calculating buyout cap hits for contracts that include signing bonuses is largely uncharted territory. However, based on information gleaned from a variety of CapGeek.com sources, it is believed signing bonuses would be paid to the player even in the event of a buyout, while the league would exclude those signing bonuses from the standard buyout calculation.

Further to that point, "actual salary" described in [1] above would exclude signing bonuses as would "remaining value" used in [3] to calculate the buyout amount.

MORE: \href{http://www.theglobeandmail.com/sports/hockey/leafs-beat/mirtle-david-clarkson-and-his-buyout-proof-contract/article17783243/?cmpid=rss1&click=dlvr.it}{David Clarkson and his buyout-proof contract (Paywall)}, The Globe and Mail, April 2, 2014

NOTE: Vincent Lecavalier's 2009-2020 contract was bought out using the 2013 CBA's compliance buyout option. In this case, there was no cap charge, but he did receive \$8 million in remaining signing bonuses on top of 2/3 of the remaining value (excluding signing bonuses) of his contract.

SEE ALSO: Can a team buy out an injured player?

\textbf{CBA Reference:} Section 50.5 (d-iii ) (Document Pages 268-271/PDF Pages 288-291)

\bigskip
\bigskip
\item
\textsc{How do Compliance Buyouts Work?}

\bigskip

During the ordinary-course buyout periods in June 2013 and June 2014, teams will be permitted two compliance buyouts (also referred to as amnesty buyouts) in total. Compliance buyouts follow the same formula as ordinary-course buyouts — either 1/3 or 2/3 of actual salary depending on age — for actual cash paid to the player, but they do not count against the cap.

Per Transition Rule 53, a compliance buyout can only be executed on a contract entered into on or before Sept. 15, 2012.

A compliance buyout can be used to eliminate the cap hit of a 35-plus contract.

For one season following a compliance buyout, the player is prohibited from rejoining the team that bought him out. According to the New York Post on June 26, the NHL notified teams that "re-signing of a player following a trade and a subsequent (compliance) buyout" would be deemed cap circumvention. However, if a team has reached its quota of two compliance buyouts, there is no defined rule preventing the team from trading a player to a team with compliance buyout slots available for the purpose of a compliance buyout.

In order for a team to execute a buyout, the player must first clear unconditional waivers. A waiver claim by another team pre-empts the buyout process. Players with no-move clauses can reject the option of waivers and proceed immediately to a buyout.

The NHL allowed teams to execute one of their two compliance buyout options prior to the 2012-13 season on players with cap hits of more than \$3 million. Two teams exercised the accelerated compliance buyout option: the Montreal Canadiens on Scott Gomez and the New York Rangers on Wade Redden. The players received full pay and counted at their full cap hit for 2012-13, after which the compliance buyout payments kick in. The \$3-million minimum applied only to accelerated compliance buyouts and does not apply to normal compliance buyouts.

The compliance buyout period coincides with the usual ordinary-course buyout period.

Ordinary-course buyouts remain available to teams in 2013 and beyond. For more information on ordinary-course buyouts, please read our separate FAQ entry.

SEE ALSO: Can a team buy out an injured player?

\bigskip
\bigskip
\item
\textsc{How do emergency goaltender tryout contracts work?}

The new CBA signed in 2013 added a feature that allows teams to sign goaltenders with professional experience to one-day professional tryout contracts in emergency conditions. Emergency conditions occur when a goaltender suffers a last-minute injury and a team is unable to dress two goaltenders because a recall cannot be effectuated in time for a game.

The previous CBA forbid goaltenders with professional experience from being signed to tryout contracts.

Goaltenders signed to one-day professional tryout (PTO) contracts cannot have a contract elsewhere, nor can they terminate an existing contract to sign the PTO. They receive US \$500 and are permitted to keep their game-worn jersey. They do not count against the cap.

One important stipulation for PTOs is that a team can’t agree to one unless they have enough cap space to recall a player with a cap hit or Averaged Amount equal to the NHL’s minimum salary. In addition, a PTO cannot be used in consecutive games.

Teams can still sign goaltenders without any professional experience to one-day amateur tryout (ATO) contracts, just as they did under the previous CBA agreed to in 2005. Players signed to ATOs receive no form of compensation and do not count against the cap.

Among other examples, the Florida Panthers signed their goaltending coach, Rob Tallas, a former NHLer, to a one-day professional tryout contract on March 3, 2013. He served as backup to Scott Clemmenson for a game against the Carolina Hurricanes. The Washington Capitals, meanwhile, signed former NCAA Division III goaltender Brett Leonhardt to a one-day amateur tryout contract on Nov. 29, 2013. He served as backup to Braden Holtby for a game against the Montreal Canadiens.

\textbf{CBA Reference:} Section 11.1 (b) and (c) on Document Page 40/PDF Page 60 and Section 16.14 on Document Page 107/PDF Page 127.

\bigskip
\bigskip
\item
\textsc{How do Entry-Level Contracts Work?}

Players younger than 25 as of Sept. 15 in the calendar year their first NHL contract is signed are subject to the terms of Article 9 of the CBA and must sign an entry-level contract (ELC). This means the amount of salary and team-paid performance bonuses is capped, depending on his draft year or the year he signs. The length of the ELC is also specified based on the player's age.

\begin{tabular}{ l | l }
\textbf{Age} & \textbf{Length} \\ 
18-21 & 3 years \\
22-23 & 2 years \\
24 & 1 year \\
25+ & No Restriction* \\
\end{tabular}

* A one-year entry-level restriction applies to drafted players from Europe aged 25 to 27.

All ELCs are two-way contracts and can be extended by as many as two seasons (see "How does a contract slide?" in this FAQ). The maximum allowable salaries for players selected in each entry draft are as follows:

\begin{tabular}{ l | l }
\textbf{Year:} & \textbf{Amount} \\
2001: & \$858,800 \\
2002: & \$900,600 \\
2003: & \$942,400 \\
2004: & \$984,200 \\
2005: & \$850,000 \\
2006: & \$850,000 \\
2007: & \$875,000 \\
2008: & \$875,000 \\
2009: & \$900,000 \\
2010: & \$900,000 \\
2011-2022: & \$925,000 \\
\end{tabular}

ELCs can contain two types of bonuses: signing bonuses and performance bonuses.

\textsc{Signing bonuses} cannot exceed 10 percent of the contract's total compensation. They are typically paid out annually.

\textsc{Performance bonuses}, or Exhibit 5 bonuses, cannot exceed \$2,850,000. These bonuses are divided into two categories: Schedule A and Schedule B.

Schedule A bonuses cannot exceed \$212,500 per individual bonus and \$850,000 in aggregate. Following is a run down.

\textsc{SKATERS}
\begin{tabular}{ p{3.5cm}  p{5cm}  p{5cm} }
 & \textit{Forwards} & \textit{Defensemen} \\
\end{tabular}

\begin{tabular}{ p{3.5cm} | p{5cm} | p{5cm} }
\textbf{Ice time (aggregate or per game)} & Player must be among top six forwards on the club (minimum 42 regular season games played by player and comparison group). An ELC may contain a provision for both aggregate and per-game ice time; however, the maximum a player may earn for ice time bonuses is \$212,500. & Player must be among top four defensemen on the club (minimum 42 regular season games played by player and comparison group). An ELC may contain a provision for both aggregate and per-game ice time; however, the maximum a player may earn for ice time bonuses is \$212,500. \\
\textbf{Goals} & 20 & 10 \\
\textbf{Assists} & 35 & 25 \\
\textbf{Points} & 60 & 40 \\
\textbf{PPG} & 0.73 (min 42 gp) & 0.49 (min 42 gp) \\
\textbf{Plus-Minus} & Among top three forwards on club (min 42 gp) & Among top three defensemen on club (min 42 gp) \\
\textbf{Blocked Shots} & — & Among top two defensemen on club (min 42 gp) \\
\textbf{All-Rookie Team} & — & — \\
\textbf{All-Star Game} & Selected to play or plays & Selected to play or plays \\
\textbf{All-Star Game} MVP & — & — \\
\multicolumn{3}{c}{}
\end{tabular}

\textsc{GOALIES}

\begin{tabular}{ p{3.6cm} | p{10cm} }
\textbf{Minutes Played} & 1,800 minimum \\
\textbf{Goals-Against Average} & GAA $\leq$ median GAA of all goalies who played 25 or more regular season games; if there is an even number of these goaltenders, the median will be the average of the two middle values. Must play in at least 25 regular season games, with at least 30 minutes in each game (including minutes replaced by an extra attacker). \\
\textbf{Save Percentage} & SV\% $\geq$ median SV\% of all goalies who played 25 or more regular season games; if there is an even number of these goaltenders, the median will be the average of the two middle values. Must play in at least 25 regular season games, with at least 30 minutes in each game (including minutes replaced by an extra attacker). \\
\textbf{Wins} & 20 (minimum 30 minutes in each game, including minutes replaced by an extra attacker) \\
\textbf{Shutouts} & Shutouts $\geq$ media number of shutouts of all goalies who played 25 or more regular season games; if there is an even number of goaltenders, the median will be the average of the two middle values. Must play in at least 30 minutes in each game (including minutes replaced by an extra attacker) \\
\textbf{All-Rookie Team} & — \\
\textbf{All-Star Game} & Selected to play or plays \\
\textbf{All-Star Game MVP} & — \\
\end{tabular}

Schedule B bonuses, with criteria and amounts outlined in the tables below, are automatically paid by the league and do not need to be written into an entry-level contract. Schedule B bonuses paid by the league do not count against the team's salary cap.
However, a player and club may agree to any compensation for the Schedule B categories listed blow, provided the maximum aggregate amount does not exceed \$2 million. That means up to \$2 million can be earned by achieving just one Schedule B bonus criteria (ie. Calder Trophy win). Schedule B bonuses paid by the team count against the team's salary cap.
Following is a list of Schedule B bonuses that are automatically paid by the league and likewise the categories that can be negotiated between the team and player.

\textsc{League-Wide Awards/Trophies}



\begin{tabular}{ l | l || l | l }
\textbf{Hart, Norris,} & & & \\
\textbf{Vezina, Selke,} & & & \\
\textbf{Richard} & & \textbf{Lady Byng} & \\
Winner: & \$250,000 & Winner: & \$150,000 \\ 
Second: & \$200,000 & Second: & \$100,000 \\
Third: & \$150,000 & Third: & \$50,000 \\
Fourth: & \$100,000 & & \\
Fifth: & \$50,000 & \textbf{Jennings:} & \\
& & Winner: & \$50,000* \\
\textbf{All-Star Team:} & & & \\
1st Team: & \$100,000 & \textbf{Calder:} & \\
2nd Team: & \$50,000 & Winner: & \$212,500 \\
 & & Third: & \$100,000 \\
\textbf{Conn Smythe:} & & Third: & \$100,000 \\
Winner: & \$250,000 & & \\
\end{tabular}

*pro-rated share if co-winner

\pagebreak

\textsc{LEAGUE PERFORMANCE BONUSES}

\textsc{Forwards}

\smallskip

\begin{tabular}{ l | l | l}
Rank & Goals, Assists, Points & PPG (min. 42 games) \\
Winner & \$150,000 &\$100,000 \\
2nd & \$140,000 & \$90,000 \\
3rd & \$130,000 & \$80,000 \\
4th & \$120,000 & \$70,000 \\ 
5th & \$110,000 & \$60,000 \\
6th & \$100,000 & \$50,000 \\
7th & \$90,000 & \$40,000 \\
8th & \$80,000 & \$30,000 \\
9th & \$70,000 & \$20,000 \\
10th & \$60,000 & \$10,000 \\
\end{tabular}

\bigskip

\textsc{Defensemen}

\smallskip

\begin{tabular}{ l | l | l }
Rank & Goals, Assists, Points & TOI, PPG (min. 42 games) \\
Winner & \$150,000 & \$100,000 \\
2nd & \$140,000 & \$90,000 \\
3rd & \$130,000 & \$80,000 \\
4th & \$120,000 & \$70,000 \\
5th & \$110,000 & \$60,000 \\
6th & \$100,000 & \$50,000 \\
7th & \$90,000 & \$40,000 \\
8th & \$80,000 & \$30,000 \\
9th & \$70,000 & \$20,000 \\
10th & \$60,000 & \$10,000 \\
\end{tabular}

\bigskip

\textsc{Goaltenders}

\smallskip

\begin{tabular}{ l | l }
Rank & GAA and SV\%$^\dagger$; wins$^\ddagger$ \\
Winner & \$150,000 \\
2nd & \$140,000 \\
3rd & \$130,000 \\
4th & \$120,000 \\
5th & \$110,000 \\
\end{tabular}

$\dagger$ Minimum 25 GP/min, 30 minu/GP

$\ddagger$ 30-minute minimum

CBA Reference: Exhibit 5 (PDF Pages: 346-351; Document Pages: 326-331).

\bigskip
\bigskip
\item
\textsc{How do Performance Bonuses count against the cap?}

\bigskip

Under the collective bargaining agreement struck in 2013, a performance bonus cushion is in place for all years of the deal, including the last one in 2021-22.

The performance bonus cushion allows teams to exceed the upper limit with performance bonuses to a maximum of 7.5 percent of the upper limit.
For example, if the upper limit is \$64,300,000, teams can exceed it by \$4,822,500 in performance bonuses. Any performances bonuses in excess of that total do not fall into the cushion and are counted as part of the team's cap payroll.

Any performance bonuses actually earned at season's end (ie. for games played, awards, all-star teams, etc.) are then added to the team's final cap payroll. The catch with the cushion is that if the bonuses earned at year's end push the team past the upper limit, the overage is carried over as a penalty the following season.

The 2009-10 Chicago Blackhawks are the most famous example of a bonus overage. Bonuses earned by the likes of Patrick Kane and Jonathan Toews following their Stanley Cup win pushed the Blackhawks' final cap payroll \$4,157,753 past the upper limit. As a result, the \$4,157,753 overage reduced their 2010-11 upper limit from \$59,400,000 to \$55,242,247 million. That, in part, resulted in the trading of Kris Versteeg, Andrew Ladd, Dustin Byfuglien, Brian Campbell, and others.


\bigskip
\bigskip
\item
\textsc{How do players injured in training camp count against the cap?}

\bigskip

Players injured in training camp can be placed on season-opening injured reserve and count against the cap in a variety of ways.

\smallskip

\textbf{Full Rate}

Players on one-way contracts or players on two-way deals who accrued 50 or more NHL games in the previous NHL season count against the cap in full while on injured reserve. Long-term injured reserve is an option for teams requiring cap relief for these players.

\textbf{Discounted Daily Rate}

Players on two-way deals who accrued 49 or fewer NHL games the previous season count at a discounted daily rate. The daily rate is calculated as follows.

Obtain a fraction by dividing the number of days the player spent on an NHL roster the previous season (including days on injured reserve) by the number of days in the previous NHL season. Multiply that fraction by the player's current cap hit to determine his daily rate.

The calculation of this daily rate for cap purposes differs slightly from that written in Section 15.6 of the CBA, which describes the calculation of the player's reduced salary rather than his reduced cap hit.
Long-term injured reserve is an option for players falling into this category. The LTI relief calculation is based on the reduced daily cap hit described above, not the full cap hit.

\textbf{Non-Counters}

Players on two-way deals who did not accrue any NHL games the previous season do not count against the cap while on season-opening injured reserve. As mentioned above, the AHL portion of the calculation outlined in Section 15.6 of the CBA pertains specifically to the calculation of the player's actual salary while on injured reserve and not his cap hit.

\textbf{CBA reference:} Section 15.6 (P. 81 and 82)

Please note accrued games are defined as games for which a player is dressed and available to play, and includes games missed due to a trade or waiver claim, as well as games not played due to illness or injury or while on a conditioning stint in the minors.

\textbf{CBA reference:} Article 21 (P. 101 and 102)

\bigskip
\bigskip
\item
\textsc{How do retained salary transactions work?}

Teams can retain a percentage of a contract’s remaining cap hit, salary and bonuses — including signing bonuses — in trades. The following basic stipulations apply:

\begin{itemize}
\item No more than 50 per cent of the salary/cap hit can be retained
\item The retained percentage must be the same for both salary and cap hit
\item The retained percentage cannot be altered from year to year
\item Salary/cap hit cannot be retained on more than three contracts per team in one season
\item The aggregate cap hits retained cannot exceed 15 per cent of the upper limit
\end{itemize}

A contract can be traded only twice where salary/cap hit is retained
In the event salary is retained a second time on the same contract, the percentage of salary/cap hit retained is applied to the full salary/cap hit and cannot alter the terms of the first retained salary transaction. Example: Devan Dubnyk was traded from Edmonton to Nashville on Jan. 15, 2014. Dubnyk's cap hit was \$3.5M and the Oilers retained 50 per cent of it (\$1.75M). Later that season, Nashville dealt Dubnyk to Montreal on March 5, 2014, retaining 25 percent of Dubnyk's contract. The percentage applied to Dubnyk's full cap hit of \$3.5M, meaning the Oilers continued to retain 50 percent of it (\$1.75M) while the Predators picked up 25 percent of it (\$875K) and the Canadiens — if Dubnyk had been on their roster — handled the remaining 25 percent (\$875K).

Regardless of whether a player with retained salary is on the acquiring team's NHL roster or loaned elsewhere (ie. the minor leagues), he continues to count against the cap of team(s) retaining his salary/cap hit. This stipulation prevents teams from circumventing the "Wade Redden Rule" for buried contracts, where without it, a team could halve a player's cap hit and potentially eliminate any cap cost while the player is "buried" in the minors. Example: Using the Dubnyk case above, the Canadiens assigned him to their AHL affiliate in Hamilton immediately after acquiring him on March 5. Although he did not count against Montreal's cap, his retained portions continued to count for Nashville (25 percent) and Edmonton (50 percent).

A team cannot reacquire a player within one calendar year of a retained salary transaction unless the player's contract expires during that time. A team also cannot reacquire in a retained salary transaction a player who was on its reserve list within the past calendar year, meaning a team can't trade a player simply to reacquire him at a reduced rate.

In the event a contract involved in a retained salary transaction is bought out, the buyout obligations — salary and cap hit — are divided between the two teams at the same percentage. For details on how retained salary transactions are treated in the event of cap advantage recapture, see the separate FAQ entry: How does cap advantage recapture work?

\bigskip
\bigskip
\item
\textsc{How do signing bonuses work?}

\bigskip

A player's NHL salary for an individual season is the sum of a base salary, and, if applicable, a signing bonus. The player's annual NHL salaries are then used to calculate the player's cap hit or annual average (AAV). In other words, signing bonuses impact cap hit.

A signing bonus can be assigned to any individual season within a contract. They can be paid up front on July 1 prior to the league year but can also doled out according to a pre-determined payment schedule that is written into the contract.

\bigskip
\bigskip
\item
\textsc{How do suspensions affect the cap?}

\bigskip

The salary of a player suspended for an on-ice incident continues to count against a team's cap. \href{http://www.broadstreethockey.com/2011/9/27/2453124/jody-shelley-suspension-salary-cap}{This was confirmed in September 2011 by Bill Daly, NHL deputy commissioner and chief legal officer, to BroadStreetHockey.com, clarifying a common misinterpretation of what is written in Exhibit 8 of the CBA in Section 5, P. 270, and in Section 50.10-c, P. 226.}

Although the CBA says salary and bonuses that are not paid to suspended players do not count against the cap, Daly clarified that the forfeited salary is still paid, albeit to the Emergency Assistant Fund. "By agreement of the NHL and NHLPA, the player's full salary continues to count against the team's cap," Daly told the website.

\bigskip
\bigskip
\item
\textsc{How do we calculate cap space?}

\bigskip

Please note the following description only applies to seasons not currently in play. Capgeek.com's in-season tracker involved more complexity.

When the bonus cushion is in effect, cap space, as referenced in features such as the cap calculator and charts, is equal to the salary cap's upper limit, plus potential performance bonuses to a maximum of 7.5 percent of the upper limit, minus cap payroll. Potential performance bonuses are the maximum earnable for the specified season. When the bonus cushion isn't in affect, cap space, as referenced in features such as the cap calculator and charts, is equal to the salary cap's upper limit minus cap payroll.

Cap payroll is equal to the sum of all cap hits on a team's roster, in addition to the sum of any buyouts and deferred bonuses (if applicable). A player's cap hit is equal to:
\[
\frac{\hbox{Total Salary} + \hbox{Total Bonuses}}{\hbox{Length (in years) of contract}}
\]

\bigskip
\bigskip
\item
\textsc{How does an Entry-Level Contract slide?}

\bigskip

If a player aged 18 or 19 signs an entry-level contract with a club (with his age calculated on Sept. 15 of the year he signed the contract) but does not play in at least 10 NHL games — regular season or playoffs — the contract will "slide" or be extended one year. The extension does not apply if the player turns 20 between Sept. 16 and Dec. 31 in the year he signed the contract.

Depending on the contract's structure, the player's cap hit can be affected either by an increase or a decrease. Players who sign at age 18 can have their contract extended (or "slide") two seasons.

While this is fairly straightforward for players returned to junior, it often leads to confusion for fans when 18- and 19-year-olds are assigned to a minor-league club. These situations are no different: the contract still slides, just as the example for \href{https://web.archive.org/web/20141018015344/http://capgeek.com/player/1578}{Tomas Tatar} illustrates.

\textbf{CBA reference:} Section 9.1 (d) (P. 23-24)

\bigskip
\bigskip
\item
\textsc{How does cap advantage recapture work?}

\bigskip

In order to punish teams for signing “back-diving” contracts under the terms of the 2005 CBA, the NHL implemented a “cap advantage recapture” rule in the 2013 CBA.

Back-diving contracts under the 2005 CBA typically had extra years with low salaries tagged on at the end with the purpose of bringing down the contract’s annual average salary/cap hit. The theory was players would retire and never play those low-salary years, meaning the player would have received more salary than the team was actually charged against the salary cap.

The cap advantage recapture rule effectively forces teams to \textbf{“pay back” any “cap advantage”} they received from these contracts, should the player retire or defect from the NHL before his contract expires. If the player fulfills his contract in full, there are no cap advantage recapture penalties.

The 2013 CBA states that any long-term contract — defined as seven years or more — entered into on or before Sept. 15, 2012, is eligible for cap advantage recapture if a team received a “cap advantage” from it at the time the player retires or defects from the NHL.

The cold, hard details of the rule follow in this FAQ entry, but if that’s not your thing, check out \href{https://web.archive.org/web/20140806210852/http://www.capgeek.com/recapture-grid}{CapGeek’s recapture grid} for a quick look at the eligible players and their penalties, or use CapGeek’s recapture calculator to test out your own scenarios.

\textbf{A “cap advantage” is accrued when the player’s actual salary exceeds his cap hit in a given year.} A negative “cap advantage” is applied to that accrual in cases where the player’s cap hit exceeds actual salary in a given season, although it should be noted that teams do not receive a credit for net negative cap advantage.

At the time of retirement/defection, \textbf{the net “cap advantage” will be “recaptured” — charged against the team’s cap in equal amounts each year until the contract expires.} This applies to any team that received a net “cap advantage” from the contract — meaning trading the player won’t make the problem go away — except for teams that dealt the player on or prior to Sept. 15, 2012.

This means that Philadelphia is not responsible for the cap advantage they received with Mike Richards and Columbus is not responsible for the cap advantage they received with Jeff Carter.

Please note, the NHL has confirmed that \textbf{contracts falling under the “over-35” rule do not qualify for cap advantage recapture.} In these cases, should the player retire before his contract expires, the team is charged with his full cap hit under the terms of the “over-35” rule. Chris Pronger is the only case where this applies.

Teams do not receive a credit for net negative cap benefit (where cap hit exceeds salary over the course of the contract prior to retirement). However, in calculating net “cap advantage,” teams do receive a credit for seasons in which cap hit exceeds salary.

Following is a step-by-step look at how these penalties are calculated for each team that received a “cap advantage” from a player’s contract, upon the player’s retirement or defection from the NHL:

\begin{itemize}
\item \textsc{Calculate Total Payment:} the total actual salary and bonuses paid to the player while with the team
\item \textsc{Calculate Total Cap Charge:} the total number of cap dollars charged against the team’s salary cap for that player
\item \textsc{Calculate Cap Advantage Recapture:} Subtract Total Cap Charge from Total Payment.
\item \textsc{Calculate Annual Cap Advantage Recapture Penalty:} Divide Cap Advantage Recapture by number of league years remaining on the contract.
\end{itemize}

\textbf{The Annual Cap Advantage Recapture Penalty applies against the team’s salary cap for each league year remaining on the contract.}

If a player retires during the season, cap advantage recapture will be calculated beginning the league year following the player’s retirement.

The first case of cap advantage recapture occurred when Ilya Kovalchuk retired from the NHL in July 2013. Kovalchuk was paid \$23 million in actual salary over the first three years of his contract while New Jersey was charged \$20 million against the salary cap. Thus, their net “cap advantage” was \$3 million, which is recaptured in equal amounts of \$250,000 per season for the remaining 12 seasons on Kovalchuk’s contract.

The league has confirmed that \textbf{both ordinary-course and compliance buyouts supersede cap benefit recapture.} However, a player must clear unconditional waivers prior to any buyout — unless he chooses to block the waiver process with a no-move clause, should he have one — and thus could be claimed by another team if the player intends to remain active. In these cases, the buyout is pre-empted and the waiving team is still subject to the recapture penalties they were trying to avoid. Meanwhile, ordinary-course buyouts executed in “back-diving” years typically come with equally stiff cap charges.

Many have speculated teams will delay the retirement of cap-advantage-recapture players and use long-term injured reserve to evade the stiff penalties. In all likelihood, that approach will only be successful if the player actually has a legitimate injury, and since the NHL went to great lengths to ensure cap advantage recapture was part of the 2013 CBA, expect it to be a watchful overseer of any such claim.

Take the example of Brian Rafalski’s retirement in 2011. He had just played a full complement of games in 2010-11 but had nagging knee and back injuries. He described himself as “banged up” but “still healthy” and made a voluntary decision to retire. If a player subject to cap advantage recapture was in a similar situation, it is unlikely a team would be successful in placing him on LTIR because that requires a player to be “unfit to play” by way of injury, illness or disability. In addition, the league has many tools at its disposal to ensure LTIR is not being abused.

\textbf{Article 50.10, Page 290:} “If the League wishes to challenge the determination of a Club physician that a Player is unfit to play for purposes of (LTIR), the League and the NHLPA shall promptly confer and jointly select a neutral physician, who shall review the Club physician's determination regarding the Player's fitness to play.”

\textbf{Article 16.11, P. 103:} “The Commissioner may take whatever steps he deems necessary to investigate the circumstances under which a Player is: (i) placed, or remains, on the Injured Reserve List, or (ii) designated Injured Non-Roster. If the Commissioner has reason to believe that the Injured Reserve List or Injured Non-Roster status has not been utilized properly by the involved Club or otherwise Circumvents any provision of this Agreement, or if he determines that the Club has used the Injured Reserve and/or Injured Non-Roster designations to evade the Active Roster limit, he may take such disciplinary action against the Club as he deems appropriate.”

Another tool at the disposal of the league, specific to the cap advantage recapture rule, is outlined in \textbf{Article 50.5, P. 267:} “The parties shall discuss in good faith the treatment of a Cap Advantage Recapture charge in the case of a retirement "orchestrated" for the sole purpose of delaying the Cap Advantage Recapture charge to the following League Year, and failing agreement, shall arbitrate the issue.”

It should be noted that contracts signed under the 2013 CBA are not eligible for cap advantage recapture because the league brought in new rules to minimize the “back-diving” issue. Limiting contract length to seven years — eight for a re-signing — and defining contract variability boundaries (See: What is the 100\% Rule?) are among them, as well as the Ilya Kovalchuk rule (See: What is the Ilya Kovalchuk Rule?, an amendment to the 2005 CBA.

\textbf{CBA Reference:} Article 50.5, P. 266-268

\bigskip
\bigskip
\item
\textsc{How does Group 6 free agency work?}

\bigskip

A Group 6 unrestricted free agent is a player with an expiring contract who meets the following criteria:

\begin{itemize}
\item Aged 25 or older (as of June 30 prior to the opening of the free agency period).
\item Has completed three or more professional seasons under NHL contract. A professional season is defined as 11 or more regular season or playoff games at ages 18 or 19 and one or more regular season and/or playoff game at ages 20 and older.
\item Has played in fewer than 80 games as a skater or 28 games as a goaltender (minimum of 30 minutes).
\end{itemize}

For players with 2012-13 counting as a professional season, transition rules to account for the lockout-shortened schedule apply and result in lower games-played requirements. The following are the adjusted games-played totals for skaters and goaltenders:

\begin{itemize}
\item Three professional seasons: 69 and 25
\item Four professional seasons: 72 and 26
\item Five professional seasons: 74 and 26
\item Six professional seasons: 75 and 27
\item Seven professional seasons: 76 and 27
\end{itemize}

Examples of how these totals are calculated follow.

Three professional seasons: $ \frac{82 + 82 + 48}{246} * 80 = 68.9 $ (round up)

Four professional seasons: $ \frac{82 + 82 + 82 + 48}{328} * 80 = 71.7 $ (round up)

\textbf{CBA reference:} Section 10.1 (c-i )(P. 28-29)

\bigskip
\bigskip
\item
\textsc{How does Long-Term Injured Reserve (LTIR) work?}

\bigskip

Teams are eligible to receive cap relief when a player is considered to have a "bona-fide long-term injury" — injuries that cause a player to miss at least 10 games and 24 days. In such cases, the team can place the player on long-term injured reserve (LTIR).

The placement of a player on LTIR does not remove the player's cap hit from the team's cap payroll — the cap hit continues to count toward the team’s cap payroll as it always did. The placement of a player on LTIR also does not provide the team with any additional cap-space savings that can be “banked” for future use while the team operates below the upper limit.

Placing a player on LTIR only provides relief if the team's cap payroll or "Averaged Club Salary" begins to exceed the upper limit. The amount of LTI relief is limited to the amount by which the team has exceeded the upper limit — less the amount of cap space the team had at the time the LTIR transaction took place. That means the amount of LTI relief is not the entire amount of the injured player's salary, although teams often make small transactions to ensure it is close.
Eligible long-term injury space is calculated on the day the player is placed on LTIR.

\textsc{Basic In-season Example}

\smallskip

The upper limit is \$70.0M. A team has a cap payroll or Averaged Club Salary of \$69.0M on the day a player with a \$4.0M cap hit is placed on LTIR. The team is now eligible to spend up to a new upper limit of \$73.0M (\$69.0M + \$4.0M).

However, had the team recalled a player with a \$750K cap hit prior to the LTIR designation, increasing its cap payroll or Averaged Club Salary to \$69.75M, it would have been eligible to spend up to a new upper limit of \$73.75M (\$69.75M + \$4.0M).

Likewise, had the team recalled two players each with \$500K cap hits prior to making the LTIR designation, increasing its cap payroll or Averaged Club Salary to exactly \$70.0M, it would have been eligible to spend to a new upper limit of \$74.0M (\$70.0M + \$4.0M).

In the above example — where the team maximized their cap space with the \$1.0M recall to get a new upper limit of \$74.0M — the team could demote the two \$500K players the day after the LTIR designation is made, reduce their cap payroll or Averaged Club Salary to \$69.0M, and have the ability to add a cap hit via trade of exactly \$5.0M.

That, of course, would only be a viable option if the player projects to be on LTIR through to the final day of the regular season. In addition, that \$5.0M figure does not increase as the season progresses, because as mentioned above, LTI cap space cannot be “banked” for future use.

These simple examples illustrate why teams in LTI situations often make what sometimes appear to be inexplicable one-day recalls and demotions. They’re manipulating their team’s cap payroll or Averaged Club Salary in order to maximize their cap space going forward.

Please note the numbers and scenarios in this example are simplified for ease of understanding.

\smallskip

\textsc{LTIR at the start of a season}

\smallskip

At the start of a season, teams facing long-term injury situations essentially have two options if they want to maximize their cap space for the rest of the year.

\textbf{Option 1:} Build the injured player into their opening-day roster and have that roster fit as close to the upper limit as possible without exceeding it, then place the player on LTIR.

By way of example, the upper limit is \$70.0M and a team’s Averaged Club Salary or cap payroll on the final day of training camp is \$71.5M, including an injured player with an Averaged Amount or cap hit of \$2M. The same day, the team demotes two players each with \$750K Averaged Amounts, reducing their Averaged Club Salary to exactly \$70.0M. Then it files its opening-day, 23-man roster to the league and places the injured player on LTIR. The team’s new upper limit is \$72.0M instead of \$70.0M for as long as the player remains on LTIR.

Had the club demoted \$2.0M in Averaged Amounts instead of the \$1.5M in the example above, their Averaged Club Salary would have been reduced to \$69.5M from \$70.0M, and their new upper limit would be just \$71.5M as opposed to \$72.0M. This highlights the importance of having the roster fit as close to the upper limit as possible before making the LTIR designation.

\textbf{Option 2:} Put the injured player on LTIR on the final day of training camp and, including the injured player’s annual average salary or cap hit, build a roster that exceeds the upper limit by an amount that is as close as possible to the injured player's annual average salary.

By way of example, the upper limit is \$70.0M and a team’s Averaged Club Salary or cap payroll on the final day of training camp is \$71.5M, including an injured player with an Averaged Amount or cap hit of \$2M. The team places the player on LTIR on the final day of training camp and is deemed to have already fully replaced the injured player, meaning the team’s new upper limit is \$71.5M instead of \$70.0M.

Had the team recalled a player with a \$500K cap hit prior to making the LTIR designation, its Averaged Club Salary would have increased to \$72.0M from \$71.5M, and likewise its new upper limit would have increased to \$72.0M from \$71.5M. This highlights the importance of a team having its Averaged Club Salary exceed the upper limit by an amount as close as possible to the injured player’s Annual Average before making the LTIR placement in order to maximize cap space.

\smallskip

\textsc{Real-World Examples}

\smallskip

Opening day of the 2013-14 featured several "real-world" examples of LTIR manoeuvring.

\smallskip

\textbf{Example 1:} The Detroit Red Wings followed Option 2. The upper limit was \$64.3M and the Red Wings’ Averaged Club Salary or cap payroll on the final day of training camp was \$66,137,878, including two injured players, Darren Helm (\$2.125M cap hit) and Patrick Eaves (\$1.2M cap hit), with an aggregate Averaged Amount of \$3.325M. Rather than simply place Helm and Eaves on LTIR, a move that would have left them with no cap space and a new upper limit of \$66,137,878, they recalled Jordan Coreau (\$792.5K) and Xavier Ouellet (\$670K), two players with an aggregate Averaged Amount totalling \$1,462,500. That increased their Averaged Club Salary to \$67,600,378 — \$3,300,378 above the upper limit or almost exactly the \$3.325M aggregate of Helm and Eaves. Only then did they place Helm and Eaves on LTIR, leaving the Red Wings with an upper limit to \$67,600,378 for the duration of the LTIR designations. The Red Wings demoted Coreau and Ouellet the next day, bringing their Averaged Club Salary back down to \$66,137,878, while leaving them with the flexibility to add \$1,462,500 in Averaged Amounts, space created by the one-day recall of Coreau and Ouellet.

\smallskip

\textbf{Example 2:} The Pittsburgh Penguins followed Option 2. The upper limit was \$64.3M and the Penguins’ Averaged Club Salary or cap payroll on the final day of training camp was \$65,990,000, which included a healthy Beau Bennett (\$990K cap hit) and an injured Tomas Vokoun (\$2M cap hit). Rather than simply place Vokoun on LTIR, a move that would have left them with no cap space and a new upper limit of \$65,990,000, they demoted Bennett and recalled Harrison Ruopp (\$650,833 cap hit) and Chris Conner (\$550K cap hit), two players with an aggregate Averaged Amount totalling \$1,200,833. That increased their Averaged Club Salary to \$66,290,833 — \$1,990,833 above the upper limit or almost exactly the \$2M cap hit of Vokoun. Only then did they place Vokoun on LTIR, leaving the Penguins with an upper limit of \$66,290,833 for the duration of the LTIR designations. The Penguins recalled Bennett and demoted Ruopp and Conner the next day, bringing their Averaged Club Salary back down to \$65,990,000, while leaving them with the flexibility to add \$300,833 in Averaged Amounts, space created by the one-day swap of Bennett for Ruopp and Conner.

\smallskip

\textbf{Example 3:} The Boston Bruins followed Option 1. The upper limit was \$64.3M and the Bruins’ Averaged Club Salary or cap payroll on the final day of training camp was \$66,086,309, including healthy Dougie Hamilton (\$894,167 cap hit) and Torey Krug (\$916,667 cap hit) — whose aggregate Averaged Amount was \$1,810,834 — and injured Marc Savard (\$4,027,143 cap hit). Rather than simply place Savard on LTIR, a move that would have left them with no cap space and a new upper limit of \$66,086,309, the Bruins demoted Hamilton and Krug on the final day of training camp, reducing their Averaged Club Salary from \$66,086,309 to \$64,275,475 — just \$24,525 under the upper limit. Only then did the Bruins place Savard on LTIR, giving them a new upper limit of \$68,302,618. The Bruins then recalled Hamilton and Krug and included them on their opening-day roster, bringing their Averaged Club Salary back up to \$66,086,309, with the flexibility to add \$2,216,309 more in Averaged Amounts, cap space that would not have existed without the one-day demotion of Hamilton and Krug. It’s important to note that the Bruins had performance bonuses exceeding the bonus cushion that reduces the figures listed in this example, but they are ignored for ease of understanding the example.

\smallskip

\textsc{Additional Notes}
Once the injured player is healthy, the team must reduce its cap payroll or Averaged Club Salary below the upper limit before the player can be removed from LTIR.

There is no formal designation of "replacement players." The league can challenge the determination of a club physician that a player is unfit to play for purposes of LTIR.

If fewer than 10 games remain in the regular season, a team can still place a player on LTIR, but the player is not eligible to return until the following season. Typically, a team would only exercise this option if the player is expected to miss the entire playoffs. If 10 or more games remain in the regular season but not 24 days, a team can still place a player on LTIR because the days requirement can extend into the playoffs. This occurred in 2013-14 with the Pittsburgh Penguins and Evgeni Malkin. Injured in a March 23 contest with 11 games and 21 days left in the regular season, Malkin was placed on LTIR on March 24. He was able to return in the playoffs on April 16.

\textbf{CBA Reference:} Section 50.10 (d) (PDF Pages: 310-314; Document Pages: 290-294)

\bigskip
\bigskip
\item
\textsc{How does the league calculate the off-season cap?}

\bigskip

Teams can exceed the salary cap's upper limit by 10 percent during the off-season. The following count toward the team's cap payroll:

\begin{itemize}
\item All players on one-way contracts, regardless of where (or if) they were playing the previous year.
\item All players on two-way contracts, in proportion to the number of days spent on an NHL roster the previous season.
\item All restricted free agents who have been extended a qualifying offer (while the offer is valid), with one-way qualifying offers counting in full and two-way qualifying offers counting as described in No. 2.
\item All restricted free agents signed to an offer sheet (such players count against the team extending the offer sheet while it is still valid).
\item All buyouts.
\end{itemize}

\smallskip

\textbf{Example:} A player with a cap hit of \$500,000 and signed to a two-way contract was on an NHL roster for 94 days in the previous season, which was 188 days long. During the off-season, he will count $\frac{\$500,000}{188} * 94 = \$250,000$

\textbf{Example:} A player with a cap hit of \$1.25 million and signed to a one-way contract was assigned to the AHL in the previous season and spent no days on an NHL roster. The player will still count \$1.25 million during the off-season.

\bigskip
\bigskip
\item 
Is Jeremy Welsh's 2011-12 cap hit a typo?

Short answer, no. Jeremy Welsh's cap hit for 2011-12 is technically \$17,945,000, the highest in NHL history. The long answer follows.

When a one-year contract containing bonuses is signed after the season has started, and is effective for the then-current season, the standard cap-hit formula is adjusted to the following:

Base = Base Salary

Days = Days in Season

Days Remaining = Days Remaining in Season

Bonus = Total Signing and Performance Bonuses

\[
\left(\left(\hbox{Base} * \frac{\hbox{Days Remaining}}{\hbox{Days}}\right) + \hbox{Bonus}\right) * \frac{\hbox{Days}}{\hbox{Days Remaining}}
\]

Examples of this in the past are entry-level contracts for Welsh with Carolina and Jesse Winchester with Ottawa in 2007-08 and Mats Sundin's one-year deal with Vancouver in 2008-09.

\bigskip

\textsc{JEREMY WELSH}

Base: \$832,500

Signing bonus: \$92,500

Performance bonus: \$0

Signed on: April 7, 2012

Days in season: 185

Days remaining: 1

\[
\left(\frac{ \$832,500 * 1 }{ 185 } + \$92,500\right) * \frac{185}{1} = \$17,945,000
\]

** NOTE: Welsh's cap hit of \$17,945,000 is compliant with \textbf{Section 50.6 (a)} of the CBA — the "20 per cent rule" — because his actual cap charge could not exceed \$97,000 based on the date he signed.

\bigskip

\textsc{JESSE WINCHESTER}

Base: \$762,500

Signing bonus: \$87,500

Performance bonus: \$25,000

Signed on: March 24, 2008

Days in season: 187

Days remaining: 13

\[
\left(\frac{ \$762,500 * 13 }{ 187 } + \$112,500 \right) *  \frac{ 187 }{ 13 } = \$2,380,769
\]

\bigskip

\textsc{MATS SUNDIN}

Base: \$1,626,500

Signing bonus: \$4,000,000

Performance bonus: \$0

Signed on: Dec. 19, 2008

Days in season: 186

Days remaining: 113

\[
\left(\frac{ \$1,626,500 * 113 }{ 186 } + \$4,000,000 \right) * \frac{ 186 }{ 113 } = \$8,210,570
\]

If the player is signing a multi-year contract under the same scenario, the standard calculation of a player's cap hit applies. Examples of this are Justin Abdelkader's entry-level deal with Detroit in 2007-08 and Antti Miettinen's deal with Tampa Bay in 2011-12.

\bigskip
\bigskip
\item
\textsc{What are the key points in the new CBA signed in 2013?}

\bigskip

Here’s a look at some preliminary details of the tentative NHL-NHLPA collective bargaining agreement. We’ve selected the details most relevant to CapGeek.com’s numbers and estimates.

To read the CBA, \href{http://www.nhlpa.com/inside-nhlpa/collective-bargaining-agreement}{please click here.}

\textsc{UPPER LIMIT (lower limit/midpoint/upper limit)}

2012-13: \$44M/\$52M/\$60M *

2013-14: \$44M/\$54.15M/\$64.3M

The lower limit must be reached without performance bonuses.
* transition rules permit teams to spend up to \$70.2M
LINK: \href{/http://www.theglobeandmail.com/sports/hockey/globe-on-hockey/how-the-nhls-salary-cap-could-hit-90-million/article7029575/}{Article by James Mirtle of the Globe and Mail (PAYWALL)}

\smallskip

\textsc{Minimum NHL Salary}

\begin{tabular}{ l | l }
2012-13: & \$525,000 \\
2013-14: & \$550,000 \\
2014-15: & \$550,000 \\
2015-16: & \$575,000 \\
2016-17: & \$575,000 \\
2017-18: & \$650,000 \\
2018-19: & \$650,000 \\
2019-20: & \$700,000 \\
2020-21: & \$700,000 \\
2021-22: & \$750,000 \\
\end{tabular}

\smallskip

\textsc{Buried Contracts (Wade Redden Rule)}
Money paid to players outside of the NHL counts against the cap.

A one-way contract counts against the cap as follows:

\[
\hbox{Cap Hit} - \left(\hbox{NHL Minimum Salary} + \$375,000\right)
\]

\textsc{Example:} If Player A with cap hit \$6.5M plays in the minors in 2013-14, he counts as follows. $ \$6,500,000 - (\$550,000 + \$375,000) = \$5,575,000 $

A two-way contract counts as follows:
\[
\hbox{Minor-League Salary} - \left(\hbox{Minimum NHL Salary} + \$375,000\right)
\]

\textsc{Example:} Player B’s two-way contract pays \$3M/\$1M in 2013-14 and counts as follows in the minors.
$\$1M - \left( \$550,000 + \$375,000 \right) = \$75,000 $

\smallskip

\textsc{Re-Entry Waivers}

Re-entry waivers have been eliminated in the new CBA.

\smallskip

\textsc{Free Agency}

Group 3 unrestricted free agency remains unchanged and begins on July 1 each year. UFAs may meet and interview with potential new clubs from the day after the entry draft (June 25 at the latest) until June 30.

\smallskip

\textsc{Contract Limits}

Maximum contract length is seven years, but extends to eight if a club is re-signing its own player. The eight-year re-signing option expires for UFAs with the opening of free agency on July 1.

\smallskip

\textsc{Contract Variability}

Front-loaded contracts where the average of salary plus bonuses over the first half of the contract exceeds the cap hit, variability rules apply. Year-to-year variability is limited to 35 per cent of the first-year compensation. If a player earns \$10M in Year 1, the contract cannot subsequently increase/decrease by more than \$3.5M from year to year. Meanwhile, the lowest year’s compensation cannot be less than 50 per cent of the highest year’s compensation. If a player earns \$10M in his highest year, he can never earn less than \$5M any other year.
The old 100 per cent rule applies to any other multi-year contract.

\smallskip

\textsc{Roster Emergencies}

When injuries or suspensions result in insufficient cap room and a shortage of skaters for more than one game, teams can — after the first game of such shortage — recall replacement players provided they have a cap hit less than or equal to the league’s minimum salary plus \$100,000, ie. \$625,000 in 2012-13.

\smallskip

\textsc{LTIR \& the 35-Plus Rule}

Both remain unchanged in the new CBA.

\textsc{Cap Advantage Recapture \& Retained Salary Transactions}

Explained Above

\bigskip
\bigskip
\item 
\textsc{What are the NHL's roster rules after the trade deadline?}

The trade deadline is set each year at 3 p.m. Eastern time on the 40th day immediately preceding the final day of the regular season. Following are rules that apply to NHL rosters on or after that date:

\begin{itemize}
\item The 23-man roster limit is lifted on the day of the trade deadline.

\item The salary cap continues to apply until the final day of the regular season, but is lifted during the playoffs.

\item Unrestricted free agents signed after the trade deadline, including those from the NCAA ranks, are not eligible to play in the NHL playoffs because they were not on the team’s 90-man reserve list on the day of the trade deadline. Example: Christian Folin with the Minnesota Wild in 2013-14.

\item Unrestricted free agents who sign on or before the trade deadline, including those from the NCAA ranks, are eligible to play in the NHL playoffs because they are on the team's 90-man reserve list on the day of the trade deadline. Example: Danny DeKeyser with the Detroit Red Wings in 2012-13.

\item Draft choices signed after the trade deadline are eligible to play in the playoffs because, by CBA definition, they were on the team's 90-man reserve list on the day of the trade deadline. Reports that such players must sign before the end of the regular season are incorrect. \textbf{Example:} Chris Kreider with the New York Rangers in 2011-12. Signed April 10, 2012, three days following the conclusion of the 2011-12 regular season on April 7, 2012. More Details: \href{https://web.archive.org/web/20150120185848/http://nypost.com/2012/04/04/college-star-kreider-could-play-for-rangers-in-playoffs/}{Via New York Post}

\item Teams are limited to just four “regular” recalls after the NHL trade deadline through the end of the regular season. Not counting toward this total are “emergency” recalls made under emergency conditions, meaning the team’s roster has been reduced below the level of two goaltenders, six defensemen and 12 forwards. In the case of an emergency recall, the player can only remain with the team for the duration of the roster emergency, after which he must be immediately returned to the club from which he was recalled. Also not counting toward this total are recalls of loaned players (ie. players on the team’s 90-man reserve list who were assigned to a minor, junior or European league) whose regular season/playoff schedules have been completed.

\item Following the conclusion of the NHL regular season, the recall limit is essentially lifted. However, a team is not permitted to have more than three players on its roster who were added by way of regular recall after the trade deadline and before the end of the regular season. Also, in the event that the team used four regular recalls between the trade deadline and the end of the season, it is permitted to retain those players on its roster following the conclusion of the regular season. Emergency recalls, explained earlier, are permitted, as are recalls of loaned players (ie. players on the team’s 90-man reserve list who were assigned to a minor, junior or European league) whose regular season/playoff schedules have been completed.

\item For an NHL-contracted player to be eligible for the AHL playoffs, the player must be on the AHL roster as of 3 p.m. Eastern time on the day of the NHL trade deadline. Teams often demote players for one day on the trade deadline to secure their AHL playoff eligibility, and sometimes, these are paper transactions of which there is no public record (ie. not reported by media or on the AHL’s official website).
\end{itemize}

An exception to this rule applies to NHL-contracted players who are signed for the current season and were on loan (regardless of league) from the NHL team at the trade deadline. 

\textbf{Example:} Teuvo Teravainen signed an NHL entry-level contract with Chicago on Aug. 22, 2013, prior to the 2013-14 season, with 2013-14 as the first season. He was loaned to Jokerit of the Finnish SM-Liiga for the 2013-14 season and was not on Rockford’s AHL roster at the 2014 NHL trade deadline. However, because he was signed for 2013-14 and on loan at that time, he was eligible to join Rockford after his SM-Liiga season ended. 

\textbf{Example:} Andreas Athanasiou signed an NHL entry-level contract with Detroit on Nov. 15, 2013, after the 2013-14 season started, with 2013-14 as the first season. He was loaned to Barrie of the OHL for the 2013-14 season and was not on Grand Rapids’ AHL roster at the 2014 NHL trade deadline. However, because he was signed for 2013-14 and on loan at that time, he was eligible to join Grand Rapids after his OHL season ended. 

\textbf{Example:} Johnny Gaudreau signed an NHL entry-level contract with Calgary on April 11, 2014, with 2013-14 as the first season. He joined the Flames immediately. Because he was not signed, and thus not on loan, at the time of the 2014 NHL trade deadline on March 5, he was not eligible to suit up for Calgary’s AHL affiliate in Abbotsford in the playoffs.

In addition, an NHL-drafted player or free agent who has signed his first NHL contract effective starting the following season is eligible to play in the AHL playoffs on an amateur tryout contract. Example: Philadelphia Flyers draft choice Shayne Gostisbehere signed an NHL entry-level contract on April 15, 2014, with 2014-15 as the first season. Since Gostisbehere was not under contract for the 2013-14 season, he was eligible to sign an amateur tryout contract to play in the 2014 AHL playoffs with the Adirondack Phantoms. Had Gostisbehere’s contract started in 2013-14, he would have been eligible to play for the Flyers in the NHL post-season but could not have been assigned to the Phantoms, similar to Johnny Gaudreau in Example 3 above.

\begin{itemize}
\item Regarding long-term injured reserve, if fewer than 10 games remain in the regular season, a team can still place a player on LTIR, but the player is not eligible to return until the following season. Typically, a team would only exercise this option if the player is expected to miss the entire playoffs. If 10 or more games remain in the regular season but not 24 days, a team can still place a player on LTIR because the days requirement can extend into the playoffs. Example: This occurred in 2013-14 with the Pittsburgh Penguins and Evgeni Malkin. Injured in a March 23 contest with 11 games and 21 days left in the regular season, Malkin was placed on LTIR on March 24. He was able to return in the playoffs on April 16.

\item Teams are technically permitted to make trades after the deadline. However, acquired players are not eligible for the NHL post-season. Historically, only minor trades have been completed prior to the conclusion of the Stanley Cup Final.
\end{itemize}

\bigskip
\bigskip
\item 
\textsc{What rules apply to contract variability and what's the 100 percent rule?}

\href{https://www.youtube.com/watch?v=yO32BvrDvAo}{Click Here for a video explanation.}

Video Transcript at the end.

Under the collective bargaining agreement struck in 2013, new contract variability rules came into play for front-loaded contracts.

For front-loaded contracts where the average of salary plus bonuses over the first half of the contract exceeds the cap hit, variability rules apply. Year-to-year variability is limited to 35 per cent of the first-year compensation. If a player earns \$10M in Year 1, the contract cannot subsequently increase/decrease by more than \$3.5M from year to year. Meanwhile, the lowest year’s compensation cannot be less than 50 per cent of the highest year’s compensation. If a player earns \$10M in his highest year, he can never earn less than \$5M any other year.

The 2005 CBA's 100 per cent rule, outlined as follows, applies to any other multi-year contract.

The difference between player salary and bonuses in each of the first two years of a contract cannot exceed the amount of the lower of the two league years. In all subsequent years, increases in player salary and bonuses from one year to another cannot exceed the amount of the lower of the first two years of the contract and decreases from one year to another may not exceed 50 percent of the lower of the first two years of the contract.

It is important to note this does not mean decrease in salary is limited to 50 percent of the prior year's salary. It's 50 percent of the lower of the first two years of the contract.

\textsc{100 PER CENT RULE EXAMPLES (min <> max listed for Year 3)}

\smallskip

\textbf{Example 1}

Year 1: \$2,000,000

Year 2: \$4,000,000

Year 3: \$3,000,000 to \$6,000,000

\smallskip
\textbf{Example 2 }

Year 1: \$1,000,000

Year 2: \$2,000,000

Year 3: \$1,500,000 to \$3,000,000

\smallskip
\textbf{Example 3 }

Year 1: \$1,000,000

Year 2: \$1,000,000

Year 3: \$500,000 to \$2,000,000

\smallskip
\textbf{Example 4 }

Year 1: \$750,000

Year 2: \$1,100,000

Year 3: \$725,000 to \$1,850,000

\smallskip
\textbf{Example 5 }

Year 1: \$2,500,000

Year 2: \$1,500,000

Year 3: \$750,000 to \$3,000,000


\textbf{CBA reference:} 50.7 (P. 282-284)

\bigskip
\bigskip
\item 
\textsc{What's a 35-Plus Contract?}

\bigskip

Players who sign multi-year contracts when they are age 35 or older (calculated on June 30 of the season the contract begins) count toward the cap under all circumstances, regardless of where (or if) the player is playing. The only cap relief is \$100,000 from the player's cap hit if he is assigned to the minors after the first year of the contract or in the event of a buyout.

The league confirmed to CapGeek.com in March 2013 that the over-35 rule "doesn't apply either in the case of regular buyouts or in the case of compliance buyouts," meaning teams can use buyouts to reduce or eliminate the cap hit. Prior to the league's clarification, CapGeek.com had operated with inaccurate information in estimating its cap totals. (See also: How do buyouts work?)

\textbf{CBA reference:} Section 50.5 (d-i-B-5) (P. 203)

\bigskip
\bigskip
\item
\textsc{What's a bonus overage?}

When the bonus cushion is in effect, teams can exceed the salary cap in performance bonuses by up to 7.5 percent of the salary cap's upper limit. Should performance bonuses actually earned at season's end push them past the salary cap's upper limit, the excess bonuses earned will be carried over as a penalty to the team's cap payroll the following season.

\bigskip
\bigskip
\item
\textsc{What's the difference between a NTC and a NMC?}

A no-trade clause means a player cannot be traded without his consent. Consent is not required for waivers for assignment to the minors. CapGeek uses the terms “NTC” and “full NTC” interchangeably.

A no-move clause means a player cannot be traded, waived for a claim by another team, or assigned to the minors without his consent. This does not protect the player from a buyout. CapGeek uses the terms “NMC” and “full NMC” interchangeably.

Limited and modified NTCs and NMCs simply mean the player's contract includes specific terms for the clauses, usually stipulating windows and/or specific seasons and/or allowing the player to specify a no-trade list of undesirable teams or a trade list of desirable teams. CapGeek uses the terms “limited” and “modified” interchangeably.

If a player has a full NMC, it travels with him even if he waives it to be traded. However, if the player is traded before his NMC takes effect, it automatically vaporizes unless — in rare circumstances — the acquiring club agrees in writing to be bound by it.

An NTC does not travel with a player if a) the player waives a full NTC or b) the player is traded within the terms of a limited or modified NTC. In addition, if the player is traded before his NTC takes effect, it vaporizes unless — in rare circumstances — the acquiring club agrees in writing to be bound by it.

It should also be noted that clauses can only cover what would have otherwise been Group 3 unrestricted free agency years.

The CBA struck in 2013 added a provision that, should the player and team agree, a NTC or NMC negotiated as part of a contract extension can take effect immediately on the player’s existing contract, providing the player was eligible for a NTC or NMC for that season (ie. it is a year that would have otherwise been a Group 3 unrestricted free agency year).

\textbf{CBA reference:} Section 11.8 (a \& b) (P. 48)


\bigskip
\bigskip
\item 
\textsc{What's the Ilya Kovalchuk rule?}

Any long-term contract — five years or more — that extends beyond the age of 40 (as of June 30 prior to the season in question) will have its traditional cap hit adjusted.

\textbf{Adjustment A}

For years of the contract up to and including the age of 40, the cap hit is calculated by dividing the total salary for those years by the number of years up to and including the 40-year-old season. For years of the contract ages 41 and beyond, the cap hit in each year is equal to the actual salary paid for that season.

\textsc{Example:} Zdeno Chara signed a seven-year, \$45.5-million deal starting in 2011 (age 34) and running through 2018 (age 41). Because he earns a \$4 million salary at age 41, his 2017-18 cap hit is \$4 million. For all other seasons, his cap hit is \$6,916,667, with his \$41.5 million in remaining salary divided over six years.

\textbf{Adjustment B}

For a contract that includes years in which the player is aged 36 through 40 (inclusive), and which averages to \$5.75 million or more for the three highest compensation seasons, the minimum single-year amount used for calculating the cap hit for ages 36 through 40 will be a minimum of \$1 million.

\textsc{Example:} Player B age 34 signs a seven-year, \$25.6-million contract starting in 2011-12 and ending at age 40 (2017-18). The cap hits are as follows:

\texttt{2011-12 (age 34): \$6m salary -------------- \$3.75m cap hit}

\texttt{2012-13 (age 35): \$6m salary -------------- \$3.75m cap hit}

\texttt{2013-14 (age 36): \$5.25m salary ----------- \$3.75m cap hit}

\texttt{2014-15 (age 37): \$5m salary -------------- \$3.75m cap hit}

\texttt{2015-16 (age 38): \$2m salary -------------- \$3.75m cap hit}

\texttt{2016-17 (age 39): \$750,000 salary --------- \$3.75m cap hit}

\texttt{2017-18 (age 40): \$600,000 salary --------- \$3.75m cap hit}

Note: In the above example, the contract averages to exactly \$5.75 million for the three highest compensation seasons, 2011-12 through 2013-14. As a result, \$1 million is used in 2016-17 and 2017-18, respectively, to calculate the cap hit, not the listed salaries of \$750,000 and \$600,000.

\textsc{Blended example:} Following is a blended example of Adjustment A and B. Player C age 34 signs a nine-year contract in 2011-12 and ending at age 42 (2019-20). The cap hits are as follows:

\texttt{2011-12 (age 34): \$6m salary -------------- \$3.75m cap hit}

\texttt{2012-13 (age 35): \$6m salary -------------- \$3.75m cap hit}

\texttt{2013-14 (age 36): \$5.25m salary ----------- \$3.75m cap hit}


\texttt{2014-15 (age 37): \$5m salary -------------- \$3.75m cap hit}

\texttt{2015-16 (age 38): \$2m salary -------------- \$3.75m cap hit}

\texttt{2016-17 (age 39): \$750,000 salary --------- \$3.75m cap hit}

\texttt{2017-18 (age 40): \$600,000 salary --------- \$3.75m cap hit}

\texttt{2018-19 (age 41): \$600,000 salary --------- \$600,000 cap hit}

\texttt{2019-20 (age 42): \$575,000 salary --------- \$575,000 cap hit}

Note: In the above example, the contract averages to exactly \$5.75 million for the three highest compensation seasons, 2011-12 through 2013-14. As a result, \$1 million is used in 2016-17 and 2017-18, respectively, to calculate the cap hit, not the listed salaries of \$750,000 and \$600,000. In addition, because the player is age 41 and 42 in the contract's final two years, those seasons are ignored for the calculation of the cap hit from 2011-12 through 2017-18. The cap hits for 2018-19 and 2019-20 match the player's salary.

The rules came into place on Sept. 4, 2010, in an attempt to cut down on long-term contracts deemed to circumvent the cap. It did not affect any long-term contracts signed prior to that date. The amendment was a direct response to the New Jersey Devils' signing of Ilya Kovalchuk to a 17-year, \$102-million contract that was rejected by the NHL on July 21, 2010, for circumventing the cap. The deal was later restructured.

\bigskip
\bigskip
\item
\textsc{What is the penalty for being over the cap?}

This question was not archived, so we don't have the capgeek answer.
The relevant section of the CBA (copied from \href{http://cdn.agilitycms.com/nhlpacom/PDF/NHL_NHLPA_2013_CBA.pdf}{here} are below and relatively readable.  Ommitted sections deal with definitions of terms and of circumvention itself:


\textbf{26.9 Violative Filings}

Any Club which files an SPC, which if approved and registered, would cause the Club's Averaged Club Salary to impermissibly exceed the Upper Limit of the Range, shall be fined a minimum of \$25,000, regardless of whether the SPC was registered and regardless of whether the terms of the SPC constitute or are the result of a Circumvention. 

\textbf{26.13 Enforcement by the System Arbitrator}

(c) In the event that the System Arbitrator finds that a Circumvention has been committed by a Player or Player Actor, the System Arbitrator may impose any or all of the following penalties and/or remedies set forth below. In the event that the System Arbitrator finds that a Circumvention has been committed by a Club or a Club Actor, the Commissioner may impose any or all of the following penalties and/or remedies set forth below: 
\begin{itemize}
\item Impose a fine of up to \$5 million in the case of a Circumvention by a Club or Club Actor, but in no circumstances shall such fine be less than \$1 million against any Club or Club Actor if such party is found to have violated Article 50 of this Agreement. If such a fine is assessed against a Club (except in the case of a financial reporting violation), that Club's Payroll Room shall also be reduced by such amount for the following League Year, and if such reduction of the Club's Payroll Room renders the Club out of compliance with the Payroll Range (i.e., the Club does not have sufficient Payroll Room to accommodate its Player commitments comprising Club Salary) for such following League Year, then the Club must take such steps as are necessary (e.g., Assignment, Buy-Out, Waivers, etc.) and as are permitted by this Agreement to ensure that the Club will be in compliance with Article 50 of this Agreement uponcommencement of the following League Year; 
\item Impose a fine against a Player of up to the lesser of \$1 million or twentyfive (25\%) percent of a Player's Paragraph 1 Salary in the case of a Circumvention by a Player or Player Actor, but in no circumstances shall such fine be below the lesser of \$250,000 or twenty-five (25\%) percent of the Player's Paragraph 1 Salary. Notwithstanding the \$1 million limitation set forth above, any additional amounts by which the Player has been unjustly enriched due to the Circumvention shall be ordered to be disgorged;
\item Direct a Club to forfeit draft picks (the number, placement, and League Year of which shall be determined in the Commissioner's sole discretion);
\item Declare a forfeiture of any NHL Game(s) determined to have been
affected by a Circumvention; 
\item Direct a Club to disclose and report to the Independent Accountants all information required by this Agreement, including, without limitation, by the provisions of Article 50;
\item Void any SPC, or any extension of an SPC, between any Player and any Club when both the Player or Player Actor and the Club or Club Actor are found to have committed such a violation with respect to such SPC or
extension; and 
\item Suspend any Club employee, Player, or Certified Agent involved in such a violation for a period of time determined in the sole discretion of the Commissioner, the System Arbitrator, or the NHLPA, respectively. 
\end{itemize}

\bigskip

\textbf{26.15 Examples of Circumvention.}

The following is a non-exhaustive list of activities that
either constitute a Circumvention under this Article 26 or from which a Circumvention may be inferred:

\begin{itemize}
\item A Club has a Club Salary that would exceed the Upper Limit, other than through the Bona-Fide Long-Term Injury/Illness Exception or the "Performance Bonus Cushion," such as by virtue of an undisclosed agreement or undisclosed payment to Players on its roster.
\item A Club has an agreement to pay money or anything else of value to a Player not expressly permitted by this Agreement, or makes such a payment to a Player. 
\item A Player enters into a sponsorship or endorsement arrangement with a local sponsor or entity with which his Club does business, in which the Player receives something of value that is disproportionate to (i.e., clearly in excess of) the fair market value of the services rendered by the Player on behalf of the sponsor or endorser.
\item A Club having access to the Averaged Amount Joint Exhibit and Averaged Club Salary Joint Exhibit and charged with the knowledge that it does not have sufficient Payroll Room to sign an SPC, as required by Article 50, proceeds to sign such SPC. Although it shall not be a Circumvention attributable or imputed to a Player if such Player signs an SPC that puts his Club out of compliance with the Payroll Range, the Certified Agent representing the Player in such circumstance, having access to the Averaged Amount Joint Exhibit and Averaged Club Salary Joint Exhibit, shall be charged with possession of the knowledge that the signing of that SPC will put the Club out of compliance with the Payroll Range.
\item A Club and a Player, during the Player's active career, agree that upon the Player's retirement, he will receive a sum of money for services to be provided to the Club after retirement.
\item A Club or Club Actor pays a Player or Player Actor for a "no-show" job, or for a job in which the payment to the Player or Player Actor clearly exceeds the fair market value of the services rendered.
\item A Club fails to report sales from one of the ticket windows in its box office.
\item A Club has an undisclosed agreement to obtain payment for tickets reported as complimentary tickets.
\item A Club has an agreement below fair market value with a Club Affiliated Entity to broadcast or otherwise present footage of the Club's NHL Games.
\item A Club fails to disclose a barter arrangement that should have been included in its
HRR Reporting Package.
\item A Club fails to report revenue generated from mascot appearances.
\end{itemize}

\bigskip
\bigskip
\item
\textsc{What are the cap implications if a player retires before his contract expires?}

This question was not archived, and this answer comes from \href{http://en.wikipedia.org/wiki/NHL_salary_cap}{Wikipedia.}

The only way to end a player's contract early is to buy it out, or have the player retire, and then only if the contract took effect prior to a player turning 35 (see the next paragraph).

The CBA also contains a 35-and-over rule, sometimes referred to as the Mogilny rule.[8] This rule states that if a player signs a multi-year deal when the player is 35 or older, starting in the second year of the contract, that amount will count towards the team's salary cap regardless of whether the player is on the active roster or not (unless the player is on long-term injured reserve); this provision remains in effect for the 2013 CBA. This is designed to keep teams from signing older players to lucrative front-loaded contracts, thus saving cap room, in which there is no expectation the player will actually play in the latter years.

The 35 rule clarified. In an email to CapGeek, the NHL informed them (prior to the 2013 CBA - and now in the 2013 CBA) - a player who signs a contract as a 35 plus can be bought out as a compliance buyout, or, as a regular buyout. As a regular buyout, only the buyout portion of that contract counts towards the cap. [9]
\end{enumerate}

\textsc{Video Transcription - Capgeek FAQ via Octagon Hockey}
My name is Andy Scott, and I'm a Certified Player Agent with Octag8n Hockey.  I'm also a lawyer, and I teach sports law at the University of Ottawa Law School.

I'm doing a series of talks dedicated to certain aspects of the 2013 Collective Bargaining Agreement.  In the next discussion, I will explain the new CBA rules on contract variability for multi-year player contracts.

However, before looking at the rules under the current CBA, it'll be important to explain the background regarding how we got to this point.  To properly understand the landscape, you need to know that as a very 
general rule, the cap hit on a player is determined by taking the overall compensation of the contract and dividing this by the number of years in the term of the contract.  Under the 2005 CBA, many NHL players were signed to long-term, front-loaded contracts which means that higher salary amounts were paid in the first years of a contract with lower salary amounts paid in the last years of the contract.  The major reason for this structure was to artificially lower the Average Annual Value (AAV) of the contract with what is referred to as "throwaway years" at the end of the contract.

\bigskip
\textsc{On screen:}
\[
\left[\hbox{\begin{tabular}{l|l|l|l}
\textsc{Year of Contract} & \textsc{NHL Salary } & \textsc{AAV} & \textsc{Cap Hit} \\
1 & \$7,000,000 & \$4,027,143 & \$4,027,143 \\
2 & \$7,000,000 & \$4,027,143 & \$4,027,143 \\
3 & \$6,500,000 & \$4,027,143 & \$4,027,143 \\
4 & \$5,000,000 & \$4,027,143 & \$4,027,143 \\
5 & \$1,500,000 & \$4,027,143 & \$4,027,143 \\
6* & \$575,000 & \$4,027,143 & \$4,027,143 \\
7* & \$575,000 & \$4,027,143 & \$4,027,143 \\
\end{tabular}}\right]
\]
*Can be argued as "throwaway years"
\bigskip

These were called "throwaway years" because it was argued that a player would actually have no intention of actually playing during those seasons, and so the NHL team could benefit from a lower cap hit on that contract, which mean that the club's overall payroll exceeded the salary cap under the CBA. [Transcription note: Holy run-on sentences Batman!]

So, this was the big issue of the time, culminating in July 2010 when Ilya Kovalchuk signed a 17 year, \$102 million deal with the New Jersey Devils that was structured to have him never sign another contract until he was 44 years old.

\bigskip
\textsc{On Screen:}
\[\left[\hbox{
\begin{tabular}{l|l|l|l}
\textsc{Year of Contract} & \textsc{NHL Salary } & \textsc{AAV} & \textsc{Cap Hit} \\
1 & \$6,000,000 & \$6,000,000 & \$6,000,000 \\
2 & \$6,000,000 & \$6,000,000 & \$6,000,000 \\
3 & \$11,500,000 & \$6,000,000 & \$6,000,000 \\
4 & \$11,500,000 & \$6,000,000 & \$6,000,000 \\
5 & \$11,500,000 & \$6,000,000 & \$6,000,000 \\
6 & \$11,500,000 & \$6,000,000 & \$6,000,000 \\
7 & \$11,500,000 & \$6,000,000 & \$6,000,000 \\
8 & \$8,500,000 & \$6,000,000 & \$6,000,000 \\
9 & \$6,500,000 & \$6,000,000 & \$6,000,000 \\
10 & \$3,500,000 & \$6,000,000 & \$6,000,000 \\
11 & \$750,000 & \$6,000,000 & \$6,000,000 \\
12 & \$550,000 & \$6,000,000 & \$6,000,000 \\
13 & \$550,000 & \$6,000,000 & \$6,000,000 \\
14 & \$550,000 & \$6,000,000 & \$6,000,000 \\
15 & \$550,000 & \$6,000,000 & \$6,000,000 \\
16 & \$550,000 & \$6,000,000 & \$6,000,000 \\
17 & \$550,000 & \$6,000,000 & \$6,000,000 \\
\end{tabular}
}\right]
\]
\indent93\% of the contract was to be paid out in the first 10 years of the 17 year deal.
\bigskip

The deal was heavily front-loaded and would have paid Ilya \$95 million in the first 10 years and only \$7 million in the last seven seasons.  It was by far the most aggressive front-loaded contract that had been filed with the league, and the NHL responded by rejecting the deal on the grounds that it circumvented the NHL's salary cap rules.  They argued that neither Kovalchuk nor the Devils expected Ilya to play out his contract.  And therefore, artificially lowered the AAV of the contract.

It is also worth noting that the NHL indicated that they were investigating the contracts of Roberto Luongo, Marc Savard, Chris Pronger, and Marian Hossa.  In response to the NHL, the NHLPA filed a grievance with respect to Kovalchuk contract, but the NHL's position was upheld in arbitration and Ilya's contract was ultimately deemed a circumvention of the CBA's salary cap rules.  Now after the decision, the NHL and the NHLPA agreed to settle the issue and passed new regulations that would govern salary cap calculations for long term contracts for the remaining term of the 2005 CBA.  As part of this settlement, the NHL dropped their investigations into the contracts of Luongo, Savard, Pronger, and Hossa, and registered a revised contract for Ilya Kovalchuk for \$100 million over 15 years.  These regulations, which won't be discussed here as they are no longer binding, paved the way for one of the NHL's top priorities in negotiations for a new CBA in 2012.  And that was closing the loophole on heavily front-loaded contracts, and passing new rules that governed the variability of an NHL contract.  Now let's discuss and deal with the new rules under the 2013 CBA.

First of all, there are two sets of rules regarding contract variability.  Front-loaded contracts and all other contracts.  Under the new CBA a contract is front-loaded if the average amount in the first half of the deal is greater than the cap hit.  I'd not that if the deal is an odd number, say five years, you just take 50\% of the compensation of the middle year and count this as part of the first half.  Also, please note that throughout this session, "compensation" will refer to salary and signing bonuses.

Let's deal with front-loaded contracts first.  For a front-loaded contract, the variability in the first three years is limited to 35\% of the first year's compensation.  Let's say a player earns \$2,000,000 in the first year of their deal.  35\% of \$2,000,000 is \$700,000, and so the year to year variability of that deal cannot be greater than \$700,000.

The second rule of front-loaded contracts is that the lowest year's compensation cannot be less than 50\% of the highest year's compensation.  Going back to our player earning \$2,000,000 in the first year of a front-loaded deal.  If that was his lowest year in terms of compensation, then he cannot earn more than \$4,000,000 in any other year of the deal.

\bigskip
\textsc{On Screen:}

\textsc{Example}
\begin{itemize}
\item First Year Compensation = \$2,000,000
\item 35\% of Compensation = \$2,000,000 * 0.35 = \$700,000
\item Therefore, the second year's compensation must be between \$1,300,000 and \$2,000,000.
\item Max Compensation, Year 2 = \$2,700,000 (Under year-to-year variability rules)
\item Max Compensation, Year 3 = \$3,400,000 (Also under year-to-year variability rules)
\item Max Compensation, Year 4 = \$4,000,000 (Under total variability limit)
\end{itemize}
\bigskip


For all contracts other than front-loaded contracts, the 100\% Rule applies.  Now this is arguably one of the most confusing and misinterpreted rules in the CBA.  But once you break it down, it is actually quite simple.  The first part of the rule involves the first two years of a multi-player contract [Note: multi-year player contract].  The variability between the first year and the second year of a contract cannot exceed the lower of those two years.

\bigskip
\textsc{On Screen:}

\[\left[\hbox{
\begin{tabular}{l l l}
 & \textsc{Valid} & \textsc{Invalid} \\
\textbf{Year} & \textbf{Compensation} & \textbf{Compensation} \\
1 & \$2,000,000 & \$2,000,000 \\
2 & \$4,000,000 & \$4,500,000 \\
\end{tabular}
}\right]
\]
\bigskip

For instance, you may have a deal which pays \$2 million in year one and \$4 million in year two.  However, you cannot have a deal that pays \$2 million in year one and \$4.5 million in year two because the variability would be greater than the lower of those two years being \$2 million.  The second part of the rule involves variability during the subsequent years of the contract, and there is a different treatment for decreases vs increases in player salary and bonuses.

Let's deal with increases first.  Any increase in compensation cannot exceed the lower of the first two years of the contract.  Going back to our example of a player earning \$2 million in year one and \$4 million in year two, the lower of those years is obviously \$2 million, and so the highest increase in compensation from one year to the next for that contract is \$2 million.  

\bigskip
\textsc{On Screen:}
\[\left[\hbox{
\begin{tabular}{l l l}
\textbf{Contract Year} & \textbf{Compensation} & \\
1 & \$2,000,000 & Maximum Increase \\
2 & \$4,000,000 & equal to lower of \\
3 & \$6,000,000 & first two years of \\
4 & \$8,000,000 & compensation
\end{tabular}}
\right]
\]
\bigskip

Now, the rules for year-to-year variability are different when it comes to decreases in compensation.  Decreases in compensation may not exceed 50\% of the lesser of the first two years of the contract. So this warrants an example.  Let's go back to our player that earns \$2 million in year one and \$4 million in year two on a non-frontloaded contract.  From one year to the next, his contract may not decrease by more than 50\% of the lowest of his first two years, and the lowest amount in the first two years was \$2 million, as we discussed.  And so 50\% of \$2 million is \$1 million, and so he cannot have a decrease in his compensation year to year by more than \$1 million.  So looking ahead to year three for that player, if you're at \$2 million in year one and \$4 million in year two, he cannot make less than \$3 million in 
year three.

\bigskip
\textsc{On Screen:}
\[
\left[\hbox{
\begin{tabular}{l l l}
\textbf{Contract Year} & \textbf{Compensation} & \\
1 & \$2,000,000 & The maximum decrease \\
2 & \$4,000,000 & allowed is \$1 million, \\ 
3 & \$3,000,000 & or half of year 1 - the \\
4 & \$2,000,000 & lower of the first two \\
\end{tabular}}\right]
\]
\bigskip

Now, another critical rule in the CBA involves the length of a player contract.  Under the old CBA, there was no express limit on the number of contracts, or the number of years sorry, for a contract, with the exception of players in the entry-level system.  Under the new CBA, the maximum term on a contract is 7 years unless a player is re-signing with the same team, in which case that term limit is 8 years. [Repeats from farther perspective]

So, to summarize our discussion today - on a front-loaded contract, the variability between years cannot be greater than 35\% of the amount of the first year's compensation.  Also, on front-loaded contracts, the lowest year's compensation may not be less than 50\% of the highest year's compensation.  With respect to contracts that are not front loaded, the 100\% rule applies.  The first part of this rule indicates that the variability between the first two years of a player contract may not exceed the compensation in the lower of those two years.  Beyond the first two years of the contract, the increase in compensation from one year to the next cannot exceed the lower of the first two years of that contract.  Again, this is for the 100\% rule.  And any decrease in compensation may not exceed 50\% of the lesser amount of the first two years of the contract.  Also, a player cannot have a term greater than seven years unless a player is re-signing with the same team, in which case they can be eight years.

Okay, so first off, I know this stuff can get convoluted and somewhat confusing, but I want to thank everyone for watching, and please to check out some of my other clips on other topics which include the entry-level system, schedule A and B bonuses, and other provisions of the CBA.  Again, my name is Andy Scott, and I'm an NHLPA Certified Agent.  If you have any questions, please contact me.  Thank you.

\bigskip

\textsc{On Screen:}

Andy Scott

Ottawa, ON

Octagon Hockey

160 Elgin Street, Suite 24

Ottawa, Ontario K2P 2P7

P: 613.238.5741

F: 613.233.7074

\smallskip

follow Andy on twitter

@Andy\_Scott15
\bigskip
\bigskip


\end{document}